Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from the biclaR project

### 2023-11-28 paper/PaperCEUS/PaperCEUS.Rmd
Update description of jittering

The mobility survey data [@IMOB] is the basis for this project and defines the baseline scenario.
Despite being conducted in the pre-pandemic period (2017), this dataset represents the most comprehensive and up-to-date information on urban mobility in Portuguese metropolitan areas (Lisbon and Porto).

We used a method for disaggregating the origins and destinations of trips between the centroids of two districts (same as "parish") to ensure that a district is not solely characterized by a single point of origin and destination for its trips.
Aggregating all trips into centroids renders the exercise less realistic, as it excludes a significant portion of short-distance trips, a prevalent characteristic of active travel [@Lovelace2022Jittering].
The OD Jittering method breaks down a single point (i.e., the centroid of an area) into multiple random points on the existing and neighboring road network, using OpenStreetMap as a reference.
This method then distributes the volume of trips within the district among the randomly generated origin-destination pairs.

Using the [`odjitter` R package](https://github.com/dabreegster/odjitter), we employed a maximum disaggregation level of 100 trips per O-D pair for this project.
Figure \ref{fig:jitter} illustrates the contrast between trip representation through the traditional method, which connects a single desire line between each district, and the presentation achieved through the randomization and disaggregation of trips between districts, specifically for the Lisbon metropolitan area.

>>>>

The mobility survey data [@IMOB] is the basis of the baseline scenario and trip rates presented in this paper.
Conducted in the pre-pandemic period (2017), this OD dataset represents the most comprehensive and up-to-date information on urban mobility in Portuguese metropolitan areas (Lisbon and Porto).

We used 'jittering' to disaggregate the OD data, resulting in a wide spatial distribution trip origins and destinations [@Lovelace2022Jittering].
The method works by sampling 'sub-points' (nodes on the transport network represented in OpenStreetMap in this case) and using these instead of a single point (typically the centroid) to represent trip origins and destinations for each zone.
This method then distributes the trips to desire lines connecting the subpoints based on a 'disaggregation threshold' which determines the maximum number of trips that can be represented by a single desire line.

Using the [`odjitter` R package](https://github.com/dabreegster/odjitter), we disaggregated the OD data into desire lines reprenting a maximum of 100 trips each.
Figure \ref{fig:jitter} illustrates the contrast between trip representation through the traditional method, which connects a single desire line between each district, and the presentation achieved through the randomization and disaggregation of trips between districts, specifically for the Lisbon metropolitan area.
